---
title: "Time Series and Econometric Analysis"

execute: 
  warning: false
  freeze: true
  cache: true
format:
  html:
    html-math-method: katex
    code-block-bg: "#E5E7E9"
    df-print: paged
    code-overflow : wrap
    code-copy : true
---

## Introduction

### Resources

-   Time series analysis by James D Hamilton - its considered to be the bible on time series analysis, pretty much covers all the theory , it doesnt have any code or pseudocode

-   Introduction to Time Series with R by Metcalfe & Coweperwait

-   Analysis of Financial Time Series by Tsay

### Aim and Motivation

-   Would be exploring time series analysis in the context of finance
-   Implementing code from the books above later 2 books mentioned above, would be using the book by Hamilton as reference
-   Aim to understand different time series models in R.
-   (Will keep on adding the aims as is go on)

## Financial Time Series and Linear Time series

### Asset Returns

::: {#def-Simple_gross_return}
## Simple gross return

<br/> <u>Arithmetic Returns</u>

$$P_t = P_{t-1}(1+R_t) \implies R_t = \frac{P_t - P_{t-1}}{P_t}$$ where $P_t$ and $R_t$ are the price and return at time $t$ respectively <br/> <u>Continuously compounded return or log return</u>

$$r_t = \ln(1+R_t)$$ <br/> <u>Dividend Payment</u>

$$r_t = \ln(P_t + D_t) -\ln(P_{t-1})$$ Where $D_t$ is dividend payment between time t-1 and time t .
:::

<br/>

::: {#def-Multiperiod_simple_return}
## Multiperiod gross return

<br/> Similiary we can define <u>multiperiod simple return</u> by $$ R_t[k] = exp{[\frac{1}{k}\sum_{j=0}^{k-1}(1+R_{t-j})]} -1$$ where exp is the exponentential function and k is total number periods the asset was held for.

<u>Log return</u> would be $$ r_t[k] = r_t +r_{t-1}+.....+r_{t-k+1}$$
:::

::: {#def-Portfolio_return}
## Portfolio Return

$$R_{p,t}= \sum_{i=1}^{N}w_i R_{it}$$

Where $p$ is the portfolio , $i$ the ith asset and $w_i$ the weight of the ith asset in the portofolio
:::

[\* Load data and header =T give the 1st row of the data file , that is the names of the cloumns of the data set]{.aside}

```{r  }
#| code-block-border-right: false
# da contains the return for 5 five stocks and indexes namely IBM ,value-weighted , equal-weighted and S&P composite index from 1970 to 2008
library(fBasics)
da = read.table("TST/d-ibm3dx7008.txt",header = T) 
dim(da) # dimensions for the data
da[1:5,]
tail(da)#first row
ibm = da[,2] # IBM simple returns

sibm = ibm*100 #Percentage simple returns 

basicStats(sibm)

s1 = skewness(sibm)
t = s1/sqrt(6/9845) #defintion for test statistic here 9845 is N 
t
pv  = 2*(1-pnorm(t)) # Calculate p-value
pv

```

```{r}
#| layout-ncol: 2
library(ggplot2)
ibm = as.data.frame(da[,1:2])
ibm["log_return"] = log(1+ibm["rtn"])
colnames(ibm)[2] <- "simple_return"
ggplot(ibm,aes(x = simple_return,colour = "Emperical"))+ggtitle("Simple return distribution") + xlim (-0.15,0.15)+geom_density()+stat_function(mapping = aes(colour = "Normal"),
fun = dnorm,args = with(ibm,c(mean =mean(simple_return),sd = sd(simple_return))))


ggplot(ibm,aes(x = log_return,colour = "Emperical"))+ggtitle("Log return distribution") + xlim (-0.15,0.15)+geom_density()+
stat_function(mapping = aes(colour = "Normal"),
fun = dnorm,args = with(ibm,c(mean = mean(log_return),sd = sd(log_return))))+
scale_x_continuous("log return distribution",limits = c(-0.15,0.15))

#stat_function(
#fun = dnorm,args = with(ibm,c(mean =mean(simple_return),sd = sd(simple_return))))+
```

```{r}

ibm1 = ibm[, 2]
plot(ibm1,type = 'l')
ibm2 = ts(ibm1,frequency = 250 ,start = c(1970,1))
plot.ts(ibm2)
acf(ibm1,lag =250,ylim= c(-0.05,0.05))
acf(ibm2,lag=250,ylim= c(-0.05,0.05))
```

### Linear Time Series

::: {#def-Stationarity}
## Stationarity

Foundation of time series analysis is Stationarity. A time series $\{r_t\}$ is said to be <u>strictly stationary</u> if the joint distribution of $(r_{t_{1}},r_{t_{2}},r_{t_{3}},....,r_{t_{k}})$ is identical to that of $(r_{t_{1+t}},r_{t_{2+t}},r_{t_{3+t}},....,r_{t_{k+t}})$ for all $t$ , where $k$ is an arbitrary positive integer and $(t_1,t_2,t_3,....,t_k)$ is collection of $k$ positive integers. Strict stationarity is very hard to verify emperically <br/> A time series is said to be <u>weakly stationary</u> if the mean of $r_t$ and the covariance between $r_t$ and $r_{t-l}$ are time invariant where $l$ is an arbitrary integer. Weak stationarity enables us to make inference cooncerning future observations

Weak stationarity is commonly studied and more practical.

The covariance $\gamma_l = Cov(r_t,r_{t-l})$ is called the <u>lag-$l$ autocovariance of $r_t$</u>. From the above defintion it follows that

-   $\gamma_0=Var(r_t)$
-   $\gamma_{-l} = \gamma_l$
:::
